---
title: "ERA_proj_code_1"
author: "Indraja Ravichandran"
date: "2024-04-29"
output: html_document
---

```{r, echo=TRUE}
library(dplyr)
set.seed(1223)

group_size_dist <- c(0, 0.3, 0.1, 0.4, 0.2)  # Percentages
group_sizes <- c(1, 2, 3, 4, 5)
arrival_time_windows <- c(15, 10, 5, 10, 15)/60  # Inter-arrival times for each hour block
simulation_hours <- seq(0, 5)  # Operating hours 5pm-10pm
current_table_mix <- c(0, 21, 3)  # Numeric vector for 2, 4, and 6-top tables
proposed_table_mix <- c(4, 17, 4)  # Numeric vector for 2, 4, and 6-top tables

# Function to simulate arrivals based on inter-arrival time distribution
simulate_arrivals <- function(inter_arrival_times, hours) {
  arrivals <- numeric()
  for (hour in hours) {
    time <- hour
    while (time < hour + 1) {
      inter_arrival <- sample(inter_arrival_times, 1, replace = TRUE)
      time <- time + inter_arrival
      if (time < hour + 1) {
        arrivals <- c(arrivals, time)
      }
    }
  }
  return(arrivals)
}
```
```{r, echo=TRUE}
simulate_night <- function(table_mix, arrival_time_windows, group_size_dist, simulation_hours) {
  tables <- rep(seq_along(table_mix), times = table_mix)
  table_states <- rep(0, length(tables))  # 0 for available, 1 for occupied
  table_release_times <- rep(0, length(tables))  # Track when each table will be free
  wait_times <- numeric()
  arrivals <- simulate_arrivals(arrival_time_windows, simulation_hours)




  # Adding variability in meal duration between 50 to 70 minutes
  meal_duration <- runif(n = length(arrivals), min = 50/60, max = 70/60)  # Meal durations in hours
  
  for (i in seq_along(arrivals)) {
    arrival <- arrivals[i]
    group_size <- sample(group_sizes, size = 1, prob = group_size_dist)
    suitable_table_index <- which(tables >= group_size & table_states == 0 & table_release_times <= arrival)
    
    if (length(suitable_table_index) > 0) {
      chosen_table <- suitable_table_index[1]
      table_states[chosen_table] <- 1
      table_release_times[chosen_table] <- arrival + meal_duration[i]
    } else {
      # Find the next available time
      if (length(table_release_times[table_release_times > arrival]) > 0) {
        wait_times <- c(wait_times, min(table_release_times[table_release_times > arrival]) - arrival)
      }
    }
  }
  
  # Convert wait times to minutes
  wait_times_in_minutes <- wait_times * 60
  
  return(list(
    avg_wait_time = mean(wait_times_in_minutes, na.rm = TRUE), 
    utilization = mean(table_states)
  ))
}

```
The simulate_night function is  key component of a simulation designed to model
the operations of a restaurant over a single evening, focusing on table occupancy 
and customer arrivals. The function takes several parameters related to the restaurant's 
seating arrangement and customer flow and sets up an environment to assess how well the 
restaurant can accommodate groups of various sizes given a specific table mix

The simulate_night function is used to simulate the restaurant's operations, focusing on seating customers and managing table turnover.

Meal Duration Variability: 
Before processing arrivals, the function generates random meal durations for each group. These durations vary between 50 and 70 minutes(mentioned in the case) and are stored in meal_duration.

Processing Arrivals: 
The function iterates through each customer arrival.
Group Size Determination: For each arrival, a group size is randomly selected based on the predefined distribution (group_size_dist).

Table Assignment: 
Then we look for an available table that can accommodate the group size and is currently unoccupied (table_states == 0) and will be available by the time of the arrival (table_release_times <= arrival).

Seating or Waiting:
If a suitable table is found, the table is marked as occupied (table_states[chosen_table] <- 1), and the time it will be released is updated based on the arrival time plus the meal duration.
If no suitable table is available, the function calculates the wait time by finding the earliest time a table will become free after the group's arrival and records this wait time.

Wait Time Conversion: After all arrivals are processed, any wait times recorded (in hours) are converted to minutes for easier interpretation.
Output: The function returns the average wait time and the utilization rate of the tables. The average wait time is calculated from wait_times_in_minutes, and utilization is calculated as the mean of table_states, representing the percentage of time that tables are occupied.
```{r, echo=TRUE}
# Function to run simulations and summarize results
run_simulation <- function(table_mix, num_simulations = 1000) {
  results <- replicate(num_simulations, simulate_night(table_mix, arrival_time_windows, group_size_dist, simulation_hours), simplify = FALSE)
  
  avg_wait_times <- sapply(results, function(res) res$avg_wait_time)
  avg_utilization <- sapply(results, function(res) res$utilization)
  
  list(
    avg_wait_time = mean(avg_wait_times),
    utilization = mean(avg_utilization) * 100
  )
}

# Run simulations for current and proposed table mixes
results_current <- run_simulation(current_table_mix)
results_proposed <- run_simulation(proposed_table_mix)

# Print the results
cat("Current Table Mix:\n")
cat(results_current$avg_wait_time, "minutes\n")
cat(results_current$utilization, "%\n\n")

cat("Proposed Table Mix:\n")
cat(results_proposed$avg_wait_time, "minutes\n")
cat(results_proposed$utilization, "%\n")


```

